---
title: "Zooplankton Data Input for LCM"
author: "Brian Mahardja"
date: "2023-04-12"
output: word_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "~/GitHub/DeltaSmelt_LCM/Salinity_Zooplankton_analysis"
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/GitHub/DeltaSmelt_LCM/Salinity_Zooplankton_analysis")
options(scipen=999)
```

Original script and dataset packages was written by Sam Bashevkin and can be found here: https://github.com/sbashevkin/FLOATDrought

# Load packages

```{r, message=FALSE, warning=FALSE}
require(conflicted)
require(MASS)
require(dplyr)
require(zooper)
require(lubridate)
require(readr)
require(tidyr)
require(ggplot2)
require(sf)
require(readxl)
require(stringr)
require(mgcv)
require(purrr)
require(deltamapr)
require(scales)

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```

# Load data

Load zooplankton model output and strata volume from Will Smith

```{r zoop modeled data, echo=T, results='hide',message=FALSE, warning=FALSE}
model_data<-read.csv("zoop sal adjustments.csv")



```

# Adjust data per Compass Memo

When 0s were introduced for baseline predictions, it resulted in infinite values for the scalars. We replaced these infinities with the maximum finite scalar calculated from model predictions for a specific action, taxon, subregion, and month (across years). When this step still yielded no finite scalar value for a given action, taxon, subregion, and month, we used the maximum finite scalar value from +/- one month.

```{r fill inf , echo=T, results='hide',message=FALSE, warning=FALSE}

##Get the maximum based on region and month

#Change infinite to NA to remove it from max calculation
model_data_maximum_region_month <- model_data
model_data_maximum_region_month[model_data_maximum_region_month == Inf] <- NA
model_data_maximum_region_month <- na.omit(model_data_maximum_region_month) 

model_data_maximum_region_month <- model_data_maximum_region_month%>% select(-year) %>% group_by(region,month,IBMR) %>% summarise_all(max) %>%
  pivot_longer(cols=starts_with("sal_"), names_to = "variable", values_to = "replaceValue_Region_Month")


##Get the maximum based on month
#Remove NA and change infinite to 0 to remove it from max calculation
model_data_maximum_month <- model_data
model_data_maximum_month[model_data_maximum_month == Inf] <- NA
model_data_maximum_month <- na.omit(model_data_maximum_month) 

model_data_maximum_month <- model_data_maximum_month %>% select(-year,-region) %>% group_by(month,IBMR) %>% summarise_all(max) %>%
  pivot_longer(cols=starts_with("sal_"), names_to = "variable", values_to = "replaceValue_Month")

##Adjust data to allow for imputation
model_data_adjusted<- model_data %>%
  pivot_longer(cols=starts_with("sal_"), names_to = "variable", values_to = "Value")
#Join data
model_data_adjusted <- model_data_adjusted %>% left_join(model_data_maximum_region_month) %>% left_join(model_data_maximum_month)
##Fill in data
model_data_adjusted$AdjustedValue <- ifelse(model_data_adjusted$Value==Inf,ifelse(is.na(model_data_adjusted$replaceValue_Region_Month),model_data_adjusted$replaceValue_Month,model_data_adjusted$replaceValue_Region_Month),model_data_adjusted$Value)

#Remove extra columns and pivot wider
model_data_adjusted <- model_data_adjusted %>% select(-c(replaceValue_Month,replaceValue_Region_Month,Value)) %>% rename(scenario = variable)

```
# Convert IBMR region to LCM region

```{r region , echo=T, results='hide',message=FALSE, warning=FALSE}

#Information from Will Smith (USFWS)
# spatial strata names: Yolo Sac SDelta EDelta LowSac LowSJ Conf SSuisunE NSuisunE Marsh SSuisunW NSuisunW
#vol<-c(100510941,144940444,225381539,65449147,89444941,259500691,163419100,153797952,76796487,121672916,107178813,184941122) # DSM2 volumes


volume_IBMR<-data.frame(region=c("Confluence","Suisun Marsh","NE Suisun","SE Suisun","NW Suisun","SW Suisun"),volume=c(163419100,121672916,76796487,153797952,184941122,107178813),LCM_region=c("West","West","West","West","West","Far West"))

#SW Suisun will be used to represent the Far West, everything else will represent West region

#Divide West region by proportion
volume_IBMR$vol_proportion<-ifelse(volume_IBMR$LCM_region=="West",volume_IBMR$volume/(163419100+121672916+76796487+153797952+184941122),1)

#Join data
model_data_adjusted <- model_data_adjusted %>% left_join(volume_IBMR)

#Break down by LCM region
model_data_adjusted_LCM <- model_data_adjusted %>% mutate(AdjustedValue_prop = AdjustedValue*vol_proportion) %>% group_by(year, month, LCM_region, scenario, IBMR) %>% 
  summarise(AdjustedValue = sum(AdjustedValue_prop))
  
```



# Use LCM zooplankton data input to calculate proportion of taxa

```{r zoop proportion, echo=T, results='hide',message=FALSE, warning=FALSE}
original_zoop_data <-read.csv("../Zooplankton_input_original/ZooMysid_74_19_df.csv")
  
zoop_data_IBMR_config <- original_zoop_data
zoop_data_IBMR_config[is.na(zoop_data_IBMR_config)] <- 0

zoop_data_IBMR_config <- zoop_data_IBMR_config %>% mutate(acartela = ACARTELA,
                                                       allcopnaup = COPNAUP+OTHCOPNAUP+EURYNAUP+SINONAUP+PDIAPNAUP,
                                                       daphnia = DAPHNIA,
                                                       eurytem = EURYTEM,
                                                       limno = LIMNOSPP+LIMNOSINE+LIMNOTET,
                                                       mysid = Hyperacanthomysis.longirostris+Neomysis.mercedis,
                                                       othcalad = OTHCALAD+SINOCAL,
                                                       othcaljuv = CALJUV+OTHCALJUV+EURYJUV+SINOCALJUV+PDIAPJUV+ASINEJUV+ACARJUV+DIAPTJUV+TORTJUV,
                                                       othclad=BOSMINA+DIAPHAN+OTHCLADO,
                                                       othcyc = AVERNAL,
                                                       pdiapfor = PDIAPFOR
                                                       ) %>% 
  select(Year,Date,Region,acartela,allcopnaup,daphnia,eurytem,limno,mysid,othcalad,othcaljuv,othclad,othcyc,pdiapfor) %>%
  mutate(Month=month(Date)) %>% select(-Date) %>%
  group_by(Year,Month,Region) %>% summarise_all(list(median)) %>%
  #calculate proportion of taxa from each sample
  mutate(total=acartela+allcopnaup+daphnia+eurytem+limno+mysid+othcalad+othcaljuv+othclad+othcyc+pdiapfor) %>%
  mutate_at(vars(acartela:pdiapfor),list(~./total)) %>% select(-total) %>%
  pivot_longer(cols=c("acartela","allcopnaup","daphnia","eurytem","limno","mysid","othcalad","othcaljuv","othclad","othcyc","pdiapfor"),
               names_to="IBMR", values_to="zoop_taxa_ratio") %>%
  rename(year=Year, month=Month, LCM_region=Region)


#Missing data to be filled with average 1995-2014 ratio for each region
zoop_data_IBMR_config_region_mean <- zoop_data_IBMR_config %>% filter(year %in% c(1995:2014)) %>% group_by(LCM_region, IBMR) %>% 
  summarise(zoop_taxa_ratio_mean = mean(zoop_taxa_ratio,na.rm=T))

model_data_for_LCM <- model_data_adjusted_LCM %>% left_join(zoop_data_IBMR_config) %>% left_join(zoop_data_IBMR_config_region_mean)

model_data_for_LCM$zoop_taxa_ratio = ifelse(is.na(model_data_for_LCM$zoop_taxa_ratio),model_data_for_LCM$zoop_taxa_ratio_mean,model_data_for_LCM$zoop_taxa_ratio)

#Other taxa from IBMR does not get used in the LCM, so we can safely remove it
model_data_for_LCM <- model_data_for_LCM %>% select(-zoop_taxa_ratio_mean) %>% filter(IBMR!="other")

#Areas and taxa without predictions are left as is (Adjust Value of 1)
model_data_for_LCM$AdjustedValue<- ifelse(is.na(model_data_for_LCM$AdjustedValue),1,model_data_for_LCM$AdjustedValue)

#Create the multiplier
model_data_for_LCM$Multiplier <- model_data_for_LCM$AdjustedValue*model_data_for_LCM$zoop_taxa_ratio

#Pivot wider to calculate the LCM taxa groups
unique(model_data_for_LCM$IBMR)

#Adjust model_data_for_LCM_adjusted to allow for easy conversion
model_data_for_LCM_adjusted <- model_data_for_LCM %>% select(-zoop_taxa_ratio,-AdjustedValue) %>%
  pivot_wider(names_from=IBMR,values_from=Multiplier) %>% mutate(
    NJ_BPUV = allcopnaup + othcaljuv,
    JA_BPUV= acartela + eurytem + limno + othcalad + othcaljuv +othcyc + pdiapfor,
    JAC_BPUV= acartela + daphnia + eurytem + limno + othcalad + othcaljuv + othclad +othcyc + pdiapfor,
    NJAC_BPUV = acartela + allcopnaup + daphnia + eurytem + limno + othcalad + othcaljuv + othclad +othcyc + pdiapfor,
    M_BPUV = mysid,
    JACM_BPUV = acartela + daphnia + eurytem + limno + mysid + othcalad + othcaljuv + othclad +othcyc + pdiapfor,
    NJACM_BPUV =acartela + allcopnaup + daphnia + eurytem + limno + mysid + othcalad + othcaljuv + othclad +othcyc + pdiapfor,
    ACM_BPUV = acartela + daphnia + eurytem + limno + mysid + othcalad + othcaljuv + othclad +othcyc + pdiapfor,
  ) %>%
  select(-acartela,-allcopnaup,-daphnia,-eurytem,-limno,-mysid,-othcalad,-othcaljuv,-othclad,-othcyc,-pdiapfor) %>%
  pivot_longer(cols=ends_with("BPUV"), names_to = "Taxa", values_to = "multiplier") %>% 
  #Rename columns
  rename(Year=year,Month=month,Region=LCM_region) %>% mutate(Month = as.character(month(Month, label = TRUE, abbr = FALSE)))


```

# Apply results according to LCM zooplankton dataset configurations

```{r conversion}
#Read data from LCM
zoop_data_LCM_raw <-read.csv("../Zooplankton_input_original/ZooMysid_74_19_df_median.csv") 
zoop_data_LCM <- zoop_data_LCM_raw %>% filter(Year %in% c(1995:2016))

#Find maximum value for each data
zoop_data_LCM_max <- zoop_data_LCM_raw %>% pivot_longer(cols=ends_with("BPUV"), names_to = "Taxa", values_to = "BPUV") %>%
  group_by(Taxa) %>% summarise(BPUV_max=max(BPUV,na.rm=T))

#Create function to expand data frame based on a column
expand.grid.df <- function(...) Reduce(function(...) merge(..., by=NULL), list(...))

#Create data frame for conversion
zoop_data_LCM_expanded <- expand.grid.df(zoop_data_LCM,unique(model_data_for_LCM_adjusted$scenario)) %>% rename(scenario=y) %>%
  pivot_longer(cols=ends_with("BPUV"), names_to = "Taxa", values_to = "BPUV") %>%
  left_join(model_data_for_LCM_adjusted)

#Multiply values by multiplier. If N/A, it's because they're in the North and South regions, which should remain unaffected.
zoop_data_LCM_scenarios <- zoop_data_LCM_expanded %>% mutate(BPUV_adjusted= ifelse(is.na(multiplier),BPUV,BPUV*multiplier)) %>%
  #Adjust all values to within the bounds of maximum amount seen
  left_join(zoop_data_LCM_max) %>% mutate(BPUV_adjusted=ifelse(BPUV_adjusted>BPUV_max, BPUV_max, BPUV_adjusted)) %>%
  select(-BPUV,-BPUV_max,-multiplier) %>%
  #Reformat data to original setup
  pivot_wider(names_from=Taxa,values_from=BPUV_adjusted) %>%
  #Remove baseline
  filter(!(scenario %in% c("sal_base_median","sal_base_l95","sal_base_u95")))

#Export the full dataset
write.csv(zoop_data_LCM_scenarios, file="ZooMysid_2022ROC_EffectsAnalysis_df_median.csv", row.names=F)

#Further aggregate and subset data per email with Will Smith (USFWS)
zoop_data_feb_mar_ACM <- zoop_data_LCM_scenarios %>% group_by(Year,Month,scenario) %>% summarise(ACM_BPUV=mean(ACM_BPUV)) %>%
  filter(Month %in% c("February","March")) %>% pivot_wider(names_from = Month, values_from = ACM_BPUV) %>%
  rename(ACM_Feb = February, ACM_Mar = March,Calendar_Year = Year)


#Export the final dataset
write.csv(zoop_data_feb_mar_ACM, file="ZooMysid_2022ROC_EffectsAnalysis_Feb_Mar_ACM.csv", row.names=F)

```

