---
title: "Delta Smelt LCME output"
author: "Brian Mahardja"
date: "2024-09-25"
output: word_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "~/GitHub/DeltaSmelt_LCM/DeltaSmeltLCME_output"
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/GitHub/DeltaSmelt_LCM/DeltaSmeltLCME_output")

options(scipen=999)

```

```{r packages, message=FALSE, warning=FALSE, echo=FALSE,error=FALSE}
require(tidyverse)
require(lubridate)
require(readxl)
require(scales)
require(conflicted)
require(knitr)
require(grid)
require(gridExtra)
require(lattice)
require(ggpubr)
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```

```{r load data, warning=FALSE, echo=FALSE, messages=FALSE,error=FALSE}
data_model <-read_xlsx("BoR Reinitiation LCME results_06-16-2025.xlsx", skip = 1, sheet="CalSim flow actions") %>% select(c(1:13)) %>%
  rename(Year = names(.)[1],Empirical=base) %>% filter(Year %in% c(1995:2015)) %>%
  rename(Alt1 = Alt1...3,
         Alt2wTUCPwoVA = Alt2v1wTUCP...4,
         Alt2woTUCPwoVA =Alt2v1woTUCP...5,
         Alt2woTUCPDeltaVA =Alt2v2noTUCP...6,
         Alt2woTUCPAllVA = Alt2v3noTUCP...7,
         Alt3=Alt3...8,
         Alt4=Alt4...9,
         Act5=Alt5...10,
         EXP1=EXP1...11,
         EXP3=EXP3...12)

wy_type<- data.frame(Year=c(1995:2016),WY_Type=c("Wet","Wet","Wet","Wet","Wet","Above Normal","Dry","Dry","Above Normal","Below Normal","Above Normal","Wet","Dry","Critically Dry","Dry","Below Normal","Wet","Below Normal","Dry","Critically Dry","Critically Dry","Below Normal")) %>%
  mutate(wet_vs_dry= ifelse(WY_Type %in% c("Wet","Above Normal"), "Wet and Above Normal years","Below Normal, Dry, or Critically Dry years"))

data_table <- data_model%>% mutate_at(vars(c(2:13)), list( ~round(., 2))) %>% mutate(Year=as.integer(Year)) %>% left_join(wy_type) %>% select(-wet_vs_dry) %>%
  rename('Sacramento Valley Water Year Index'=WY_Type)

data_fig <- data_model %>% gather("Alternatives","lambda",2:13) %>% mutate(Year=as.numeric(Year),Alternatives=as.factor(Alternatives))

# Create table with caption
knitr::kable(data_table , padding=0, caption="Table 3. Predicted population growth rate (lambda) for each cohort year by alternatives.")


data_wy <- data_fig %>% left_join(wy_type) %>% 
  group_by(wet_vs_dry,Alternatives) %>% summarise(geometric_mean_lambda=exp(mean(log(lambda)))) %>%
  spread(Alternatives,geometric_mean_lambda) %>% rename(Category=wet_vs_dry)

data_time <- data_fig %>% mutate(TimePeriod=ifelse(Year < 2006,"1995-2005","2006-2015")) %>% 
  group_by(TimePeriod,Alternatives) %>% summarise(geometric_mean_lambda=exp(mean(log(lambda))))%>%
  spread(Alternatives,geometric_mean_lambda) %>% rename(Category=TimePeriod)

data_sum <- data_fig %>%
  group_by(Alternatives) %>% summarise(geometric_mean_lambda=exp(mean(log(lambda)))) %>%
  mutate(Category="1995-2015") %>% spread(Alternatives,geometric_mean_lambda)

data_combine_sum <- bind_rows(data_sum,data_wy,data_time) %>% mutate_at(vars(c(2:13)), list( ~round(., 2)))

# Create table with caption
knitr::kable(data_combine_sum , padding=0, caption="Table 4. Geometric mean of population growth rate (lambda) for each alternative.") 

```

```{r barplot_sum, echo=FALSE, messages=FALSE, dpi=300, fig.width=6, fig.height=8}

safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#882255", "red", "#999933", "#6699CC", "#888888")

data_bar <- data_combine_sum %>% gather("Alternatives","lambda",2:13) %>%
  filter(Category=="1995-2015") %>% mutate(Alternatives=as.factor(Alternatives))


plot_bar <- ggplot(data_bar,aes(x=Alternatives,y=lambda,fill=Alternatives)) +  
  geom_col() +
  theme_bw() +
  theme(plot.title=element_text(size=13), 
                 axis.text.x=element_text(size=9, color="black",angle=45,hjust=1), 
                 axis.text.y = element_text(size=8, color="black"), 
                 axis.title.x = element_blank(),
                 axis.title.y = element_text(size = 8, angle = 90),
                 strip.text = element_text(size = 7),legend.text=element_text(size=6),
                 legend.position = "none") + guides(fill="none") + 
  geom_hline(yintercept = 1, col = "red",linetype=2) +
  scale_fill_manual(values = safe_colorblind_palette,name="")+
  ylab("1995-2015 Geometric Mean of Population Growth Rate")


data_sum_2 <- data_sum %>% gather("Alternatives","lambda",2:12) %>% mutate(lambda_relative = (lambda-0.9811579)/0.9811579) %>% filter(Alternatives!="NAA")

plot_bar2 <- ggplot(data_sum_2,aes(x=Alternatives,y=lambda_relative,fill=Alternatives)) +  
  geom_col() +
  theme_bw() +
  theme(plot.title=element_text(size=13), 
                 axis.text.x=element_text(size=9, color="black",angle=45,hjust=1), 
                 axis.text.y = element_text(size=8, color="black"), 
                 axis.title.x = element_blank(),
                 axis.title.y = element_text(size = 8, angle = 90),
                 strip.text = element_text(size = 7),legend.text=element_text(size=6),
                 legend.position = "none") + guides(fill="none") +  
  scale_fill_manual(values = safe_colorblind_palette,name="")+
  ylab("Mean porportional change in lambda vs NAA")

grid.arrange(plot_bar, plot_bar2, nrow = 2)



```
Figure 12. Top: Bar plot demonstrating the geometric mean of population growth rate (lambda) from 1995 to 2015 for the various alternatives as seen in Table 4. Bottom: Bar plot demonstrating the relative difference in geometric mean of population growth rate (1995-2015) for each alternative compared to the no action alternative 
```{r eis, echo=FALSE, messages=FALSE, dpi=300, fig.width=6, fig.height=4}

data_bar_eis <- data_bar %>% dplyr::filter(!Alternatives %in% c("EXP1","EXP3"))

plot_bar_EIS <- ggplot(data_bar_eis,aes(x=Alternatives,y=lambda,fill=Alternatives)) +  
  geom_col() +
  geom_hline(yintercept = 1, col = "red",linetype=2) +
  theme_bw() +
  theme(plot.title=element_text(size=13), 
                 axis.text.x=element_text(size=9, color="black",angle=45,hjust=1), 
                 axis.text.y = element_text(size=8, color="black"), 
                 axis.title.x = element_blank(),
                 axis.title.y = element_text(size = 8, angle = 90),
                 strip.text = element_text(size = 7),legend.text=element_text(size=6),
                 legend.position = "none") + guides(fill="none") +  
  scale_fill_manual(values = safe_colorblind_palette,name="")+
  ylab("1995-2015 Geometric Mean of Population Growth Rate")
plot_bar_EIS

```

```{r lambda plot, echo=FALSE, messages=FALSE, dpi=300, fig.width=6, fig.height=8}

plot_lambda <- ggplot(data_fig) +  
  geom_line(aes(x=Year, lambda, colour=Alternatives),linewidth=1) + 
  theme_bw() +
  scale_x_continuous(breaks = c(1995:2015))+
  theme(plot.title=element_text(size=13), 
                 axis.text.x=element_text(size=9, color="black",angle=45,hjust=1), 
                 axis.text.y = element_text(size=8, color="black"), 
                 axis.title.x = element_blank(),
                 axis.title.y = element_text(size = 9, angle = 90),legend.text=element_text(size=6),
                 strip.text = element_text(size = 7),
                 legend.position = "top") + 
  ylab("Population Growth Rate")+
  scale_colour_manual(values = safe_colorblind_palette,name="")+
  guides(color = guide_legend(reverse = TRUE,nrow=3,byrow=TRUE))

data_naa <- data_model %>% select(Year, NAA) %>% mutate(Year=as.numeric(Year))

data_fig_percent <- data_model %>% select(-Empirical) %>% gather("Alternatives","lambda",2:11) %>%
  mutate(Year=as.numeric(Year),Alternatives=as.factor(Alternatives)) %>% left_join(data_naa) %>%
  ##Divide by NAA
  mutate(lambda_relative_to_NAA = lambda/NAA*100)

safe_colorblind_palette2 <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99",  "#882255","#999933", "red", "#6699CC", "#888888")

plot_relative <- ggplot(data_fig_percent) +  
  geom_line(aes(x=Year, lambda_relative_to_NAA, colour=Alternatives),linewidth=1) + 
  theme_bw() + scale_y_continuous(breaks = c(0,50,100,150,200,250,300), limits= c(0,305))+
  scale_x_continuous(breaks = c(1995:2015))+
  theme(plot.title=element_text(size=13), 
                 axis.text.x=element_text(size=9, color="black",angle=45,hjust=1), 
                 axis.text.y = element_text(size=8, color="black"), 
                 axis.title.x = element_blank(),
                 axis.title.y = element_text(size = 9, angle = 90),
                 strip.text = element_text(size = 7),legend.text=element_text(size=6),
                 legend.position = "top") + 
  ylab("% change in population growth rate from NAA")+
  scale_colour_manual(values = safe_colorblind_palette2,name="")+
  guides(color = guide_legend(reverse = TRUE,nrow=3,byrow=TRUE))

grid.arrange(plot_lambda, plot_relative, nrow = 2)

```

Figure 11. Top: Line plot of population growth rate (lambda) across alternatives as seen in Table 3. Bottom: Line plot showing % change calculated as lambda rate for a given alternative divided by estimated population growth rate for NAA (no action alternative). Note the color change for NAA in the bottom figure.

```{r data input, echo=FALSE, messages=FALSE, dpi=300, fig.width=9, fig.height=13}
#Code to check data input and compare with actual data
data_input_ROC <- read.csv(file.path("../Salinity_Zooplankton_analysis/FlowZoopData_2022ROC_EffectsAnalysis_CohortYear_2024-09-25.csv")) %>%
  # Change outflow values to total volume in acre foot
  mutate(Outflow_Jun0Aug0 = Outflow_Jun0Aug0 *1.983)

#Change names per Julia Taylor's email
unique(data_input_ROC$scenario)
data_input_ROC <- data_input_ROC %>% rename(scenario_original=scenario) %>% mutate(scenario = case_when(
  scenario_original == "Alt5" ~ "Act5",
  scenario_original == "Alt1" ~ "Alt1",
  scenario_original == "Alt2v1wTUCP" ~ "Alt2wTUCPwoVA",
  scenario_original == "Alt2v1woTUCP" ~ "Alt2woTUCPwoVA",
  scenario_original == "Alt2v2noTUCP" ~ "Alt2woTUCPDeltaVA",
  scenario_original == "Alt2v3noTUCP" ~ "Alt2woTUCPAllVA",
  scenario_original == "Alt3" ~ "Alt3",
  scenario_original == "Alt4" ~ "Alt4",
  scenario_original == "EXP1" ~ "EXP1",
  scenario_original == "EXP3" ~ "EXP3",
  scenario_original == "NAA" ~ "NAA"
))

data_input_ROC$scenario_original<-NULL
str(data_input_ROC)
data_input_ROC <- data_input_ROC %>% 
  dplyr::select(Cohort_Year, scenario, ACM_BPUV_Feb1Feb1, ACM_BPUV_Mar1Mar1, Outflow_Jun0Aug0,OMR_AprMar,OMR_Jun,OMR_DecJan, OMR_Feb, OMR_Mar)

data_lcme_original <-read_xlsx(file.path("../Zooplankton_input_original/LCME covariates.xlsx"),sheet="Best model covariates")

data_lcme_subset <- data_lcme_original %>% select(any_of(colnames(data_input_ROC))) %>% 
  mutate(scenario="Empirical")
  
data_input_ROC<- data_input_ROC%>%
  bind_rows(data_lcme_subset) %>% mutate(scenario=as.factor(scenario),linesize=ifelse(scenario=="Empirical",1.5,1))

#Write a function to produce plot
datainput_plot <- function(df, titletext,ytitletext,xtitletext,parameter) {
ggplot(data=df) + geom_line(aes(x=Cohort_Year,y=parameter,group=scenario, colour=scenario,linewidth=linesize)) +
      labs(title = titletext,y = ytitletext, x = xtitletext)+
    scale_linewidth(range = c(1,1.2))+
    scale_x_continuous(breaks = c(1995:2016))+
    theme(plot.title=element_text(size=13), 
                 axis.text.x=element_text(size=9, color="black",angle=45,hjust=1), 
                 axis.text.y = element_text(size=8, color="black"), 
                 axis.title.x = element_blank(),
                 axis.title.y = element_text(size = 9, angle = 90),
                 strip.text = element_text(size = 7),
                 legend.position = "top") + 
  scale_colour_manual(values = safe_colorblind_palette,name="")+
  guides(color = guide_legend(reverse = TRUE,nrow=3,byrow=TRUE),linewidth="none")
}

 # scale_linewidth_manual(values = c(Alt1=1,Alt2v1woTUCP=1,Alt2v1wTUCP=1,Alt2v2noTUCP=1,Alt2v3noTUCP=1,Alt3=1,Alt4=1,EXP1=1,EXP3=1,Empirical=2,NAA=1)) +
 

plot1<-datainput_plot(df=data_input_ROC, titletext="",ytitletext = "Sum of Delta Outflow (June-August) in acre-foot",xtitletext="Cohort Year",parameter=data_input_ROC$Outflow_Jun0Aug0)+ theme(legend.position = "none")

plot2<-datainput_plot(df=data_input_ROC, titletext="",ytitletext = "February Food Biomass per Volume Index",xtitletext="Cohort Year",parameter=data_input_ROC$ACM_BPUV_Feb1Feb1)+ theme(legend.position = "none")

plot3<-datainput_plot(df=data_input_ROC, titletext="",ytitletext = "March Food Biomass per Volume Index",xtitletext="Cohort Year",parameter=data_input_ROC$ACM_BPUV_Mar1Mar1)+ theme(legend.position = "none") +
      theme(axis.title.x = element_text(size = 11)) + xlab("Cohort Year")

# Add legend
plot8<-datainput_plot(df=data_input_ROC, titletext="",ytitletext = "March daily average OMR in cfs",xtitletext="Cohort Year",parameter=data_input_ROC$OMR_Mar) +
  guides(color = guide_legend(reverse = TRUE,nrow=5,byrow=TRUE))

leg <- get_legend(plot8)

grid.arrange(plot1, plot2, plot3,as_ggplot(leg), ncol = 1)
```
Figure 3. Outflow and prey metric data based on CalSim3 data and salinity-zooplankton model relative to the original dataset used to build the Delta Smelt LCME (labeled as "Empirical"): June-August sum of Delta outflow, February and March prey metric (biomass per volume) data composed of copepod adults, cladocerans, and mysids 

```{r plot 4,echo=FALSE, messages=FALSE, dpi=300, fig.width=12, fig.height=16}
plot4<-datainput_plot(df=data_input_ROC, titletext="",ytitletext = "April-May daily average OMR in cfs",xtitletext="Cohort Year",parameter=data_input_ROC$OMR_AprMar) + theme(legend.position = "none")

plot5<-datainput_plot(df=data_input_ROC, titletext="",ytitletext = "June daily average OMR in cfs",xtitletext="Cohort Year",parameter=data_input_ROC$OMR_Jun) + theme(legend.position = "none")

plot6<-datainput_plot(df=data_input_ROC, titletext="",ytitletext = "December-January daily average OMR in cfs",xtitletext="Cohort Year",parameter=data_input_ROC$OMR_DecJan) + theme(legend.position = "none")

plot7<-datainput_plot(df=data_input_ROC, titletext="",ytitletext = "February daily average OMR in cfs",xtitletext="Cohort Year",parameter=data_input_ROC$OMR_Feb) + theme(legend.position = "none")



plot8<-datainput_plot(df=data_input_ROC, titletext="",ytitletext = "March daily average OMR in cfs",xtitletext="Cohort Year",parameter=data_input_ROC$OMR_Mar) +  theme(
                 legend.position = "none",legend.text=element_text(size=7))

grid.arrange(plot4, plot5, plot6, plot7, plot8,as_ggplot(leg), ncol = 2)

```

Figure 4. Daily average OMR flow data produced from CalSim3 relative to the original dataset used to build the Delta Smelt LCME (labeled as "Empirical")


```{r boxplot,echo=FALSE, messages=FALSE, dpi=300, fig.width=6, fig.height=8}
data_input_ROC_reconfig <- data_input_ROC %>% select(-linesize) %>% gather("Parameter","Value",3:10) %>%
  mutate(Parameter_name = case_when(Parameter == "ACM_BPUV_Feb1Feb1" ~ "February Food Biomass per Volume Index" ,
                                    Parameter == "ACM_BPUV_Mar1Mar1" ~ "March Food Biomass per Volume Index" ,
                                    Parameter == "Outflow_Jun0Aug0" ~ "Sum of Delta Outflow (June-August)" ,
                                    Parameter == "OMR_AprMar" ~  "April-May daily average OMR",
                                    Parameter == "OMR_Jun" ~  "June daily average OMR",
                                    Parameter == "OMR_DecJan" ~ "December-January daily average OMR",
                                    Parameter == "OMR_Feb" ~  "February daily average OMR",
                                    Parameter == "OMR_Mar" ~ "March daily average OMR" ))


#Create figure
plot_parameter <- ggplot2::ggplot(data=data_input_ROC_reconfig, ggplot2::aes(x=scenario, y=Value))+ facet_wrap(~ Parameter_name, ncol=2, scales="free")+
  ggplot2::theme_bw()+
  geom_boxplot(outlier.colour="black", outlier.shape=16,
             outlier.size=2, notch=FALSE)+
  ggplot2::theme(plot.title=element_text(size=9), 
                 axis.text.x=element_text(size=9, color="black",angle=45,hjust=1), 
                 axis.text.y = element_text(size=8, color="black"), 
                 axis.title.x = element_blank(), 
                 axis.title.y = element_text(size = 9, angle = 90),
                 strip.text = element_text(size = 7),
                 strip.background = element_rect(size=0.3))+
  ggplot2::ylab("Values")
plot_parameter

```
```{r barplot_param,echo=FALSE, messages=FALSE, dpi=300, fig.width=6, fig.height=8}
data_input_ROC_reconfig_sum <- data_input_ROC_reconfig %>% group_by(scenario,Parameter,Parameter_name) %>% summarise(Value=mean(Value))


#Create figure
plot_bar_param <- ggplot2::ggplot(data=data_input_ROC_reconfig_sum, ggplot2::aes(x=scenario, y=Value, fill=scenario))+ facet_wrap(~ Parameter_name, ncol=2, scales="free")+
  ggplot2::theme_bw()+
  geom_col() +
  ggplot2::theme(plot.title=element_text(size=9), 
                 axis.text.x=element_text(size=9, color="black",angle=45,hjust=1), 
                 axis.text.y = element_text(size=8, color="black"), 
                 axis.title.x = element_blank(), 
                 axis.title.y = element_text(size = 9, angle = 90),
                 strip.text = element_text(size = 7),
                 strip.background = element_rect(size=0.3))+
  ggplot2::ylab("Values")+ scale_fill_manual(values = safe_colorblind_palette,name="")
plot_bar_param


```


```{r barplot_wet,echo=FALSE, messages=FALSE, dpi=300, fig.width=6, fig.height=8}
data_input_table <- data_input_ROC %>% left_join(wy_type %>% rename(Cohort_Year=Year))  %>% 
  select(-Cohort_Year,-linesize,-WY_Type,-wet_vs_dry) %>%
  group_by(scenario) %>% summarise_each(funs(mean))

data_input_table_wy <- data_input_ROC %>% left_join(wy_type %>% rename(Cohort_Year=Year))  %>% 
  select(-Cohort_Year,-linesize,-WY_Type) %>%
  group_by(scenario,wet_vs_dry) %>% summarise_each(funs(mean))

data_input_barplot_wet<- data_input_table_wy %>% filter(wet_vs_dry=="Wet and Above Normal years") %>%
  gather("Parameter","Value",3:10) %>%
  mutate(Parameter_name = case_when(Parameter == "ACM_BPUV_Feb1Feb1" ~ "February Food Biomass " ,
                                    Parameter == "ACM_BPUV_Mar1Mar1" ~ "March Food Biomass" ,
                                    Parameter == "Outflow_Jun0Aug0" ~ "June-August Delta Outflow" ,
                                    Parameter == "OMR_AprMar" ~  "April-May OMR",
                                    Parameter == "OMR_Jun" ~  "June OMR",
                                    Parameter == "OMR_DecJan" ~ "December-January OMR",
                                    Parameter == "OMR_Feb" ~  "February OMR",
                                    Parameter == "OMR_Mar" ~ "March OMR" ))

#Create figure
plot_bar_param_wet <- ggplot2::ggplot(data=data_input_barplot_wet, ggplot2::aes(x=scenario, y=Value, fill=scenario))+ facet_wrap(~ Parameter_name, ncol=2, scales="free")+
  ggplot2::theme_bw()+
  geom_col() +
  ggplot2::theme(plot.title=element_text(size=9), 
                 axis.text.x=element_text(size=9, color="black",angle=45,hjust=1), 
                 axis.text.y = element_text(size=8, color="black"), 
                 axis.title.x = element_blank(), 
                 axis.title.y = element_text(size = 9, angle = 90),
                 strip.text = element_text(size = 7),
                 strip.background = element_rect(size=0.3))+
  ggplot2::ylab("Values")+ scale_fill_manual(values = safe_colorblind_palette,name="")
plot_bar_param_wet


```
Figure Y1. Mean value of variables used in the LCME for Wet and Above Normal year types. Note that cohort year was matched with the water year that the cohort was born in (e.g., cohort year 1995 = water year 1995).

```{r barplot_dry,echo=FALSE, messages=FALSE, dpi=300, fig.width=6, fig.height=8}

data_input_barplot_dry<- data_input_table_wy %>% filter(wet_vs_dry=="Below Normal, Dry, or Critically Dry years") %>%
  gather("Parameter","Value",3:10) %>%
  mutate(Parameter_name = case_when(Parameter == "ACM_BPUV_Feb1Feb1" ~ "February Food Biomass " ,
                                    Parameter == "ACM_BPUV_Mar1Mar1" ~ "March Food Biomass" ,
                                    Parameter == "Outflow_Jun0Aug0" ~ "June-August Delta Outflow" ,
                                    Parameter == "OMR_AprMar" ~  "April-May OMR",
                                    Parameter == "OMR_Jun" ~  "June OMR",
                                    Parameter == "OMR_DecJan" ~ "December-January OMR",
                                    Parameter == "OMR_Feb" ~  "February OMR",
                                    Parameter == "OMR_Mar" ~ "March OMR" ))

#Create figure
plot_bar_param_dry <- ggplot2::ggplot(data=data_input_barplot_dry, ggplot2::aes(x=scenario, y=Value, fill=scenario))+ facet_wrap(~ Parameter_name, ncol=2, scales="free")+
  ggplot2::theme_bw()+
  geom_col() +
  ggplot2::theme(plot.title=element_text(size=9), 
                 axis.text.x=element_text(size=9, color="black",angle=45,hjust=1), 
                 axis.text.y = element_text(size=8, color="black"), 
                 axis.title.x = element_blank(), 
                 axis.title.y = element_text(size = 9, angle = 90),
                 strip.text = element_text(size = 7),
                 strip.background = element_rect(size=0.3))+
  ggplot2::ylab("Values")+ scale_fill_manual(values = safe_colorblind_palette,name="")
plot_bar_param_dry

```

Figure Y2. Mean value of variables used in the LCME for Below Normal, Dry, or Critically Dry year  Note that cohort year was matched with the water year that the cohort was born in (e.g., cohort year 1995 = water year 1995).